---
title: "Week 4 FLS"
author: "Eric Graham"
date: "2024-09-17"
output: 
  powerpoint_presentation:
    slide_level: 2
    reference_doc: presentation.pptx
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, results = "hide")
library(dplyr)
library(ggplot2)
library(ggthemes)
library(scales)
```

# 1: Quick Quiz Questions

## QQQ 1

1. D
2. C
3. B
4. A
5. E

## QQQ 2

A, C, F, H

## QQQ 3

A t-test on the difference in means between the log income for the two groups would be more powerful. Welch's t-test is designed to account for samples with high variance, so it sacrifices some power to do so. In this case, since the log-transformed data produced normal distributions with similar variance and spread, we can use a pooled t-test which is more powerful.

## QQQ 4

# 2: Reaction Time Experiment

## Power Test in R

```{r, echo=TRUE}
power.t.test(n = 250, delta = 5, sd = 25, type = "paired", alternative = "two.sided", sig.level = .05)
```

## Answers

1. The probability that our test will make a mistake and reject the null hypothesis if it's true is alpha, or significance level, which is given as 0.05.
2. This would be a Type I error.
3. If the true mean is 107, then the effect size is 5.
4. The probability that our test will make the correct decision and reject the null hypothesis is 0.8829952.
5. "Power" is the statistical term for this value.
6. ![ATTACH DRAWING]()

# 3: Recall the Reaction Time Experiment

## Power Test in R

```{r, echo=TRUE}
power.t.test(n = 41, delta = 0.07, sd = 0.2, type = "paired", alternative = "two.sided", sig.level = .05)
```

## Answers

1. The probability that our test will make a mistake and reject the null hypothesis if it's true is alpha, or significance level, which is given as 0.05.
2. The power would be .5898
![ATTACH DRAWING]()
3. 

# 4: Recall the Creativity Study

# 5: Takeaways

# 6: Questions

