---
title: "Week  FLS"
author: "Eric Graham"
date: "2024- - "
output: 
  powerpoint_presentation:
    slide_level: 2
    reference_doc: presentation.pptx
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, results = "hide")
library(dplyr)
library(ggplot2)
library(ggthemes)
library(scales)
```

For Live Session: Calculate the correlation coefficient for the data Professor Sadler included in the previous lightboard lecture, and then plot the data. Then give a brief explanation as to what the value that you calculate for the correlation coefficient means for the data set. You may do this on paper and / or in Excel.  Either way, please add it to your FLS PowerPoint. 

For Live Session
Use SAS and R to plot the data and find the linear correlation coefficient r for the speed and MPG data. Please add a screenshot of the output and your code to the FLS PowerPoint deck.

```{r}


study_time = data.frame(score = c(34, 56, 45, 70, 55, 68, 67, 79, 45, 89, 95, 78, 94), hours = c(1, 1, 2, 2, 2, 3, 4, 4, 4, 6, 7, 7, 8))
cor(study_time)
cor.test(study_time$hours, study_time$score)
fit = lm(score~hours, data = study_time)
summary(fit)
```

```{r}
SwimTemp = data.frame(swimsuits = c(5,10,15,17,20,30,37,38,45,55), temperature = c(64,62,68,70,71,80,82,85,87,92))
fit = lm(swimsuits~temperature, data = SwimTemp)
summary(fit)
```

https://www.rossmanchance.com/applets/2021/regshuffle/regshuffle.htm

# Querstion

wtf is bivariate normal distribution, where is the bell?
Speed MPG

20   24

30   28

40   30

50  28

60  24