---
title: "Week  FLS"
author: "Eric Graham"
date: "2024- - "
output: 
  powerpoint_presentation:
    slide_level: 2
    reference_doc: presentation.pptx
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(investr)
```

# QQQ

## Answers

1. a
2. c (there could be heteroscedasticity!)
3. d

# Movie Prediction/Calibration

## 1: Plot data / six step test for linear relationship

With such a small sample, it is hard to identify a clear linear relationship, but it does appear that higher budgets track with higher gross earning. Fortunately, we can test the data to see if a linear relationship exists.

```{r, echo = FALSE}
MovieData = data.frame(
    Budget = c(62, 90, 50, 35, 200, 100, 90),
    Gross = c(65, 64, 48, 57, 601, 146, 47)
)

ggplot(MovieData, aes(x = Budget, y = Gross)) +
  geom_point(shape = 19) +
  labs(title = "Movie Budget vs Gross (in Millions)", x = "Budget", y = "Gross") +
  theme_few()
```

---

### Step 1: Problem Statement

We want to test whether there is a linear relationship between movie budget and gross earnings. Our null hypothesis is that slope of the line of best fit for our data (meaning that there is no linear relationship).

$$H_0: \hat{\beta}_1 \text{ (slope)}= 0$$

$$H_0: \hat{\beta}_1 \text{ (slope)}\neq 0$$

---

### Step 2: Critical Value

R shows the critical value to be +/- 2.57 for this data (degrees of freedom = n-2 = 5).

```{r}
qt(.05/2, 5)
```

---

### Step 3, 4, 5: Test Statistic, P-Value, and Decision

R's linear model function provides the test statistic and p-value for this data. The test statistic of 5.478 is much higher than the critical value, leading to a p-value of .00276 and a decision to reject the null hypothesis.

```{r}
fit = lm(Gross ~ Budget, data=MovieData)
summary(fit)
```

---

### Step 6: Conclusion and Confidence Interval

There is strong statistical evidence (p-value of .00276) to suggest that the slope of the line of regression for the relationship between budget and earnings is not equal to zero. We are 95% certain that the slope of the line of regression is between 1.84 and 5.1, which would indicate a linear relationship between a movie's budget and its earnings. 

```{r}
confint(fit)
```

## 2 and 3: Common Standard Deviation and Confidence Interval for Slope

As shown in the summary of the linear model above, the estimated common standard deviation is 0.6338.

As shown in the conclusion above, the confidence interval for the slope is between 1.84 and 5.1.

## 4: Prediction for Gross Budget of $95 million

By adding a new value of 95 for "budget" we can predict that the estimated earnings for a $95 million budget are $165.71 million, with a 95% prediction interval of $83.52 million to $247.89 million (I notice that the prediction interval for an individual point is narrower than the confidence interval for the mean of the slope!).

```{r}
newpoint=data.frame(Budget=95)
prediction_95 = predict(fit, newdata=newpoint, interval="confidence", level=0.95)
prediction_95
```

## 5: Prediction for Gross Budget of $200 million

By adding a new value of 200 for "budget" we can predict that the estimated earnings for a $95 million budget are $530.28 million, with a 95% prediction interval of $237.4 million to $823.15 million.

```{r}
newpoint=data.frame(Budget=200)
prediction_200 = predict(fit, newdata=newpoint, interval="prediction", level=0.95)
prediction_200
```

## 6: Plots for 95% Confidence and Prediction Intervals

```{r, echo = FALSE}
fit_95 = prediction_95[1, "fit"]
lwr_95 = prediction_95[1, "lwr"]
upr_95 = prediction_95[1, "upr"]

ggplot(MovieData, aes(x = Budget, y = Gross)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  geom_point(aes(x = 95, y = fit_95), color = "red", size = 3) +
  geom_point(aes(x = 95, y = lwr_95), color = "red", size = 2) +
  geom_point(aes(x = 95, y = upr_95), color = "red", size = 2) +
  geom_errorbar(aes(x = 95, ymin = lwr_95, ymax = upr_95), color = "red") +
  geom_text(aes(x = 95, y = fit_95, label = "Est. = 165.7"), hjust = 1.1, vjust = .1, color = "red") +
  geom_text(aes(x = 95, y = lwr_95, label = "Lower = 83.5"), hjust = -.1, color = "red") +
  geom_text(aes(x = 95, y = upr_95, label = "Upper = 247.9"), hjust = 1.1, color = "red") +
  theme_few() +
  labs(title = "Linear Model with Prediction/CI for $95 millon Budget",
       x = "Budget",
       y = "Gross")
```

---

```{r, echo = FALSE}
fit_200 = prediction_200[1, "fit"]
lwr_200 = prediction_200[1, "lwr"]
upr_200 = prediction_200[1, "upr"]

ggplot(MovieData, aes(x = Budget, y = Gross)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  geom_point(aes(x = 200, y = fit_200), color = "red", size = 3) +
  geom_point(aes(x = 200, y = lwr_200), color = "red", size = 2) +
  geom_point(aes(x = 200, y = upr_200), color = "red", size = 2) +
  geom_errorbar(aes(x = 200, ymin = lwr_200, ymax = upr_200), color = "red") +
  geom_text(aes(x = 200, y = fit_200, label = "Est. = 530.3"), vjust = -.25, hjust = 1.1, color = "red") +
  geom_text(aes(x = 200, y = lwr_200, label = "Lower = 237.4"), hjust = 1.1, color = "red") +
  geom_text(aes(x = 200, y = upr_200, label = "Upper = 823.2"), hjust = 1.1, color = "red") +
  theme_few() +
  labs(title = "Linear Model with Prediction/CI for $200 millon Budget",
       x = "Budget",
       y = "Gross")
```

## 7: Calibration Interval for Gross $210 million

By using calibrate() from the investr package, we can see that the estimated budget to ensure a gross of $210 million is $107.76 million, with a 95% confidence interval of $40.65 million to $174.86 million.

```{r}
calibrate(fit, y0=210, interval="Wald", mean.response=FALSE, limit=FALSE)
```

# Takeaways and Questions

## Takeaways

1. Confidence intervals for the slope of a regression line can be used to assess the precision of a linear model (based on width or narrowness). This allows us to make predictions about the range of possible means for an explanatory value at a particular level of significance.
2. Prediction intervals can be used to provide a range of predictions for a specific explanatory value; they are wider than confidence intervals because estimating a specific value introduces more uncertainty than estimating a mean. 
3. Calibration is the process of reverse-engineering a range of plausible explanatory (x) values from a given response variable (y). 

## Questions

1. I feel pretty good about this, I hope I'm right!
2. Not stat-related, but there must be an easier way to create labels (or even arrows) for specific points in ggplot.